# Obtain data and perform QC using Seurat{#seuratqc}


## Download data

The PBMC multiome dataset is available from [10x genomics](https://support.10xgenomics.com/single-cell-multiome-atac-gex/datasets/1.0.0/pbmc_granulocyte_sorted_10k). The website contains output files from CelL Ranger ARC. Please go to the above website and make sure you download the following files to your data folder:

* Filtered feature barcode matrix (HDF5)
* ATAC Per fragment information file (TSV.GZ)
* ATAC Per fragment information index (TSV.GZ index)

In Terminal window:
```{bash eval=FALSE}
cd ~/pbmc_data
# download data
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_filtered_feature_bc_matrix.h5
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_atac_fragments.tsv.gz
wget https://cf.10xgenomics.com/samples/cell-arc/1.0.0/pbmc_granulocyte_sorted_10k/pbmc_granulocyte_sorted_10k_atac_fragments.tsv.gz.tbi

# check files
ls
```

## Load data in R and construct Seurat object

For the following processing and analysis steps, we are going to use `Seurat` and its sister package `Signac`. `Seurat` is a popular package in R that provides users with well-curated functions and workflows. Seurat was first developed for the clustering of scRNA-seq data, but with continuing updates in the last few years, this package has become a popular tool for QC, analysis, and exploration of scRNA-seq data as well. For more information on `Seurat`, see the [Seurat website](https://satijalab.org/seurat/) from the Satija Lab, which has very nice documentation and links to the Satija Lab [publications](https://satijalab.org/publications) as well as detailed [tutorials and vignettes](https://satijalab.org/seurat/vignettes.html).

In this notebook, we will use the most recent version of Seurat, version 5.

We first load all required libraries in this tutorial.
```{r eval=TRUE,results=FALSE,warning=FALSE,message=FALSE}
library(Seurat)
library(SeuratData)
library(Signac)
library(BSgenome.Hsapiens.UCSC.hg38) 
library(EnsDb.Hsapiens.v86)
library(ggplot2)
library(cowplot)
library(hdf5r)
library(dplyr)
library(patchwork)
library(clustree)
library(TFBSTools)
library(JASPAR2020)
library(BiocParallel)
library(motifmatchr)
```

Also set working directory to the folder where you saved your downloaded datasets.
```{r eval=FALSE}
# check current working directory
getwd()

# set working directory
setwd("/dfs6/pub/your_user_name")
getwd() # check again
```


Next, we will read in the `.h5` file into R and construct Seurat Object with both scRNA-seq and scATAC-seq assays.
```{r eval=F,results='markup',warning=FALSE,message=FALSE}
# the 10x hdf5 file contains both data types. 
inputdata.10x <- Read10X_h5("pbmc_granulocyte_sorted_10k_filtered_feature_bc_matrix.h5")
# extract RNA and ATAC data
rna_counts <- inputdata.10x$`Gene Expression`
atac_counts <- inputdata.10x$Peaks

## if reading files from "filtered_feature_bc_matrix"
# inputdata.10x <- Read10X(data.dir = "filtered_feature_bc_matrix")

```

```{r eval=F,results='markup',warning=FALSE,message=FALSE}
# Create Seurat object
pbmc <- CreateSeuratObject(counts = rna_counts,project="pbmc")

# Now add in the ATAC-seq data
# we'll only use peaks in standard chromosomes
grange.counts <- StringToGRanges(rownames(atac_counts), sep = c(":", "-"))
grange.use <- seqnames(grange.counts) %in% standardChromosomes(grange.counts)
atac_counts <- atac_counts[as.vector(grange.use), ]
annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotations) <- 'UCSC'
genome(annotations) <- "hg38"

frag.file <- "pbmc_granulocyte_sorted_10k_atac_fragments.tsv.gz"
chrom_assay <- CreateChromatinAssay(
   counts = atac_counts,
   sep = c(":", "-"),
   genome = 'hg38',
   fragments = frag.file,
   min.cells = 10,
   annotation = annotations
 )
pbmc[["ATAC"]] <- chrom_assay

# check data
pbmc
```

Add something about seurat object structure
Explore Seurat Objects using `@` or `$`
```{r eval=FALSE}
# check available assays in data
pbmc@active.assay

#check a single assay
pbmc[["RNA"]]
pbmc[["ATAC"]]

# check elements within each assay

pbmc[["RNA"]]$counts

# check metadata
pbmc@meta.data
```

Saving data periodically to preserve Seurat results. Output data can be large.

```{r eval=F, echo=F}
saveRDS(pbmc,file="pbmc.rds")
```

## Perform quality control using both assays
```{r eval=T,echo=F}
pbmc<-readRDS("../pbmc_data/pbmc.rds")
```

Using `DefaultAssay()` to specify the assay you want to work on.

```{r eval=FALSE,warning=FALSE,message=FALSE}
# calculate percentage of mitochondrial genes in each cell
DefaultAssay(pbmc) <- "RNA"
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
DefaultAssay(pbmc) <- "ATAC"
pbmc <- NucleosomeSignal(pbmc)
pbmc <- TSSEnrichment(pbmc)

DensityScatter(pbmc, x = 'nCount_ATAC', y = 'TSS.enrichment', log_x = TRUE, quantiles = TRUE)

# generate violin plot
VlnPlot(pbmc, features = c( "nCount_RNA", "nFeature_RNA", "percent.mt"), ncol = 3,
  log = TRUE, pt.size = 0) + NoLegend()
VlnPlot(pbmc, features = c("nCount_ATAC", "TSS.enrichment", "nucleosome_signal"), ncol = 3,
  log = TRUE, pt.size = 0) + NoLegend()
# change pt.size to show scatter p
```

```{r eval=T,fig.cap='Voilin plot for QC', out.width='100%', fig.asp=.75, fig.align='center'}
# show plot
VlnPlot(pbmc, features = c( "nCount_RNA", "nFeature_RNA", "percent.mt"), ncol = 3,
  log = TRUE, pt.size = 0) + NoLegend()
VlnPlot(pbmc, features = c("nCount_ATAC", "TSS.enrichment", "nucleosome_signal"), ncol = 3,
  log = TRUE, pt.size = 0) + NoLegend()
```



```{r eval=FALSE,warning=FALSE,message=FALSE}
# play with different filtering cutoffs to see how number of cells change
pbmc <- subset(
  x = pbmc,
  subset = nCount_ATAC < 7e4 &
    nCount_ATAC > 5e3 &
    nCount_RNA < 25000 &
    nCount_RNA > 1000 &
    percent.mt < 25 &
    nucleosome_signal < 2 &
    TSS.enrichment > 1
)

pbmc

```

Saving data periodically to preserve Seurat results. Output data can be large.

```{r eval=F}
saveRDS(pbmc,file="pbmc_qc.rds")
```
*******
