# Data integration
The PBMC multiome data was built in the `SeuratData` package and was stored in a SeuratObject already, so we can download the data and directly start analyzing. This data only needs to be loaded once, and R will skip this step the next time you load it.

```{r eval=T, echo=F}
pbmc<-readRDS("../pbmc_data/pbmc_anno_new.rds")
```

```{r eval=FALSE}
# load both modalities
pbmc.rna <-pbmc[["RNA"]]
pbmc.atac <- pbmc[["ATAC"]]

```

Identifying anchors between scRNA-seq and scATAC-seq datasets
In order to identify ‘anchors’ between scRNA-seq and scATAC-seq experiments, we first generate a rough estimate of the transcriptional activity of each gene by quantifying ATAC-seq counts in the 2 kb-upstream region and gene body, using the GeneActivity() function in the Signac package. The ensuing gene activity scores from the scATAC-seq data are then used as input for canonical correlation analysis, along with the gene expression quantifications from scRNA-seq. We perform this quantification for all genes identified as being highly variable from the scRNA-seq dataset.

```{r}
# quantify gene activity
gene.activities <- GeneActivity(pbmc.atac, features = VariableFeatures(pbmc.rna))

# add gene activities as a new assay
pbmc.atac[["ACTIVITY"]] <- CreateAssayObject(counts = gene.activities)

# normalize gene activities
DefaultAssay(pbmc.atac) <- "ACTIVITY"
pbmc.atac <- NormalizeData(pbmc.atac)
pbmc.atac <- ScaleData(pbmc.atac, features = rownames(pbmc.atac))

# Identify anchors
transfer.anchors <- FindTransferAnchors(reference = pbmc.rna, query = pbmc.atac, features = VariableFeatures(object = pbmc.rna),
    reference.assay = "RNA", query.assay = "ACTIVITY", reduction = "cca")
```

Annotate scATAC-seq cells via label transfer
After identifying anchors, we can transfer annotations from the scRNA-seq dataset onto the scATAC-seq cells. The annotations are stored in the seurat_annotations field, and are provided as input to the refdata parameter. The output will contain a matrix with predictions and confidence scores for each ATAC-seq cell.

```{r}
celltype.predictions <- TransferData(anchorset = transfer.anchors, refdata = pbmc.rna$seurat_annotations,
    weight.reduction = pbmc.atac[["lsi"]], dims = 2:30)

pbmc.atac <- AddMetaData(pbmc.atac, metadata = celltype.predictions)

```

After performing transfer, the ATAC-seq cells have predicted annotations (transferred from the scRNA-seq dataset) stored in the predicted.id field. Since these cells were measured with the multiome kit, we also have a ground-truth annotation that can be used for evaluation. You can see that the predicted and actual annotations are extremely similar.

```{r}
pbmc.atac$annotation_correct <- pbmc.atac$predicted.id == pbmc.atac$seurat_annotations
p1 <- DimPlot(pbmc.atac, group.by = "predicted.id", label = TRUE) + NoLegend() + ggtitle("Predicted annotation")
p2 <- DimPlot(pbmc.atac, group.by = "seurat_annotations", label = TRUE) + NoLegend() + ggtitle("Ground-truth annotation")
p1 | p2

```

In this example, the annotation for an scATAC-seq profile is correctly predicted via scRNA-seq integration ~90% of the time. In addition, the prediction.score.max field quantifies the uncertainty associated with our predicted annotations. We can see that cells that are correctly annotated are typically associated with high prediction scores (>90%), while cells that are incorrectly annotated are associated with sharply lower prediction scores (<50%). Incorrect assignments also tend to reflect closely related cell types (i.e. Intermediate vs. Naive B cells).

```{r}
predictions <- table(pbmc.atac$seurat_annotations, pbmc.atac$predicted.id)
predictions <- predictions/rowSums(predictions)  # normalize for number of cells in each cell type
predictions <- as.data.frame(predictions)
p1 <- ggplot(predictions, aes(Var1, Var2, fill = Freq)) + geom_tile() + scale_fill_gradient(name = "Fraction of cells",
    low = "#ffffc8", high = "#7d0025") + xlab("Cell type annotation (RNA)") + ylab("Predicted cell type label (ATAC)") +
    theme_cowplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

correct <- length(which(pbmc.atac$seurat_annotations == pbmc.atac$predicted.id))
incorrect <- length(which(pbmc.atac$seurat_annotations != pbmc.atac$predicted.id))
data <- FetchData(pbmc.atac, vars = c("prediction.score.max", "annotation_correct"))
p2 <- ggplot(data, aes(prediction.score.max, fill = annotation_correct, colour = annotation_correct)) +
    geom_density(alpha = 0.5) + theme_cowplot() + scale_fill_discrete(name = "Annotation Correct",
    labels = c(paste0("FALSE (n = ", incorrect, ")"), paste0("TRUE (n = ", correct, ")"))) + scale_color_discrete(name = "Annotation Correct",
    labels = c(paste0("FALSE (n = ", incorrect, ")"), paste0("TRUE (n = ", correct, ")"))) + xlab("Prediction Score")
p1 + p2

```



